### 2017-02-01 manuscript.rmd
bib fix
added null simulation

```{r preliminaries, echo=FALSE, message=F, warning=F}
	library( treeio )
	library( tidyverse )
	library( magrittr )
	library( stringr )
	library( parallel )
	library( ggtree )
	library( ape )
	library( digest )
	library( hutan )  # From https://github.com/caseywdunn/hutan
	library( gridExtra )

@@

		# Calculate the contrasts
		p = pic( nhx@data$Tau[ is.tip.nhx( nhx ) ], nhx@phylo, var.contrasts=TRUE )

@@

		# Make a new human readable column for plots that records speciation/ duplication
		nodes_all$Event = NA
		nodes_all$Event[ nodes_all$D == "Y" ] = "Duplication"
		nodes_all$Event[ nodes_all$D == "N" ] = "Speciation"

@@

	# Annotate each tree by joining the corresponding expression data to the 
	# @data object. This maps the expression data to the tree tips
	gene_trees_annotated = mclapply(
		gene_trees,
		function( tree ){
			tree@data %<>% left_join( expression, by = c( "G" = "Ensembl.Gene.ID" ) )
			return( tree )
		}, 
		mc.cores=cores
	)

@@

		# Merge data of this mrca to the node
		D %<>% left_join(
			nhx@data %>%
				select( D, node, node_age, label, pic ),
			by = c( "mrca" = "node")

@@

	pairwise_summary$Event = NA
	pairwise_summary$Event[ pairwise_summary$D == "N" ] = "Speciation"
	pairwise_summary$Event[ pairwise_summary$D == "Y" ] = "Duplication"

@@

				nhx@data$Event = NA
				nhx@data$Event[ nhx@data$D == "Y" ] = "Duplication"
				nhx@data$Event[ nhx@data$D == "N" ] = "Speciation"

@@

>>>>

```{r preliminaries, echo=FALSE, message=F, warning=F}
	library( treeio )
	library( tidyverse )
	library( magrittr )
	library( stringr )
	library( parallel )
	library( ggtree )
	library( ape )
	library( digest )
	library( hutan )  # From https://github.com/caseywdunn/hutan
	library( gridExtra )
	library( geiger )
	library( phytools )

@@

		tau = nhx@data$Tau[ is.tip.nhx( nhx ) ]

		if( any( is.na(tau) ) ){
			stop( "A trait value is NA." )
		}

		# Calculate the contrasts
		p = pic( tau, nhx@phylo, var.contrasts=TRUE )

@@

@@

	# Annotate each tree by joining the corresponding expression data to the 
	# @data object. This maps the expression data to the tree tips
	gene_trees_annotated = mclapply(
		gene_trees,
		function( tree ){
			tree@data %<>% left_join( expression, by = c( "G" = "Ensembl.Gene.ID" ) )

			tree@data$Event = NA
			tree@data$Event[ tree@data$D == "N" ] = "Speciation"
			tree@data$Event[ tree@data$D == "Y" ] = "Duplication"

			return( tree )
		}, 
		mc.cores=cores
	)

@@

		# Merge data of this mrca to the node
		D %<>% left_join(
			nhx@data %>%
				select( D, Event, node, node_age, label, pic ),
			by = c( "mrca" = "node")

@@

@@

@@

#### Simulations

```{r simulations_null, cache=TRUE, echo=FALSE, fig.cap="XX_PIC_VAR", warning=FALSE, message=FALSE}

	sim_tau = function( nhx ) {
		tau_original = nhx@data$Tau [ 1:length(nhx@phylo$tip.label) ]
		names( tau_original ) = nhx@phylo$tip.label
		brownian_model = fitContinuous( nhx@phylo, tau_original )
		x = fastBM( 
				nhx@phylo, 
				a=brownian_model$opt$z0, 
				sig2=brownian_model$opt$sigsq, 
				bounds=c(-Inf,Inf) 
			) %>% abs()
		x[ x > 1 ] = 1
		x[ is.nan( x ) ] = 1
		names( x ) = NULL
		nhx@data$Tau = c( x, rep( NA, nrow( nhx@data ) - length (x ) ) )
		return( nhx )
	}

	# Replace the actual Tau values with simulated Tau values, using parameters 
	# fitted from each tree
	gene_trees_sim_null = lapply( gene_trees_calibrated[1:1000], sim_tau )

	# Perform contrasts 
	gene_trees_sim_null %<>% add_pics_to_trees()

	# Summarize results
	nodes_sim_null_contrasts = gene_trees_sim_null %>% summarize_contrasts()

	# Plot
	g_phylo_null = 
		nodes_sim_null_contrasts %>% 
		ggplot( aes( x=pic, y=..density.., col=Event ) ) + 
			geom_freqpoly( binwidth=.001, position="identity" ) +
			xlab( "Phylogenetic independent \ncontrasts (PIC)" ) +
			ylab( "Density" ) +
			xlim( 0, 0.075 ) + 
			theme_classic() + 
			ggtitle( "(b) Phylogenetic result under null simulation" )

	pairwise_sim_null_summary = 
		mclapply( gene_trees_sim_null, get_pairwise_summary, mc.cores=cores ) %>% 
		bind_rows()

	# Calculate some summary statistics 
	ortholog_r_sim_null = 
		pairwise_sim_null_summary %>% 
		filter( D=="N" ) %>% 
		select( tau_a, tau_b ) %>% 
		cor( ) %>% 
		.[ 1, 2 ]

	paralog_r_sim_null = 
		pairwise_sim_null_summary %>% 
		filter( D=="Y" ) %>% 
		select( tau_a, tau_b ) %>% 
		cor( ) %>% 
		.[ 1, 2 ]

	g_pair_null = 
		pairwise_sim_null_summary %>% 
		filter( species_a == "homo_sapiens" | species_b == "homo_sapiens" ) %>% 
		filter( ! is.na( D ) ) %>%
		group_by( Event, label ) %>% 
		summarise( cor.test(  tau_a, tau_b )$estimate, n=n() ) %>% 
		rename( rho=`cor.test(tau_a, tau_b)$estimate` ) %>%
		rename( clade = label ) %>% 
		left_join( calibration_times_all, clade=clade ) %>%
		ggplot() + 
			geom_point( aes( x=age, y=rho, col=Event, size=n ) ) +
			theme_classic() +
			ggtitle( "(a) Pairwise result under null simulation" )

```{r simulations_plot, cache=TRUE, echo=FALSE, fig.cap="XX_PIC_VAR", warning=FALSE, message=FALSE}

	grid.arrange(
		g_pair_null,
		g_phylo_null,
		ncol = 2
	)

> Figure XXSim

We first simulated an analysis under the null model that tau evolves at the same rate following duplication and speciation events. Under the null model, pairwise approaches are very similar as for the observed data (Figure XXSim a), with higher rho for orthologs (`r ortholog_r_sim_null`) than paralogs (`r paralog_r_sim_null`).  Phylogenetic approaches do show similar divergence following speciation and duplication (Figure XXSim b) and does not reject the null hypothesis that the rates differ (Wilcoxon p=`r wilcox_oc( nodes_sim_null_contrasts )`).
